Rule-Based Preordering on Multiple Syntactic Levels in Statistical Machine Translation

### 2014-09-24 Paper/paper.tex
.

\address{Institute for Anthropomatics \\
Karlsruhe Institute of Technoloy, Germany \\
{\small \tt utcur@student.kit.edu, yuqi.zhang@kit.edu}
}
%
\begin{document}
\maketitle
%
\begin{abstract}
We propose a novel data-driven rule-based preordering approach, which uses the information of syntax tree and word alignment to reorder the words in source sentences before decoding in a phrase-based SMT system between English and Chinese. The preorering algorithm extracts reordering patterns from multiple levels of the syntax trees of the training data and applies the rules on the source sentences in a similar manner. We've conducted experiments in English-to-Chinese and Chinese-to-English translation directions. Our results show that the approach has led to improved translation quality both when it was applied separately on the baseline or when it was combined with some other reordering approaches. As our reordering approach was used alone, it showed an improvement of $1.61$ in BLEU score in the English-to-Chinese translation direction and an improvement of $2.16$ in BLEU score in the Chinese-to-English translation direction, in comparison with the baseline, which used no preordering at all. As our preordering approach were combined with the short rule \cite{short}, long rule \cite{long} and tree rule \cite{tree} based preordering approaches, it showed an improvement of $0.43$ in BLEU score in the English-to-Chinese translation direction and an improvment of $0.3$ in BLEU score in the Chinese-to-English translation direction. Through the translations that used our preordering approach, we've also found improvements in syntactic structure.
\end{abstract}
%
\section{Introduction}

@@

Rottman and Vogel 2007 \cite{short} introduced the idea of extracting reordering rules from the POS tag sequences of training data and use them for reordering. Niehues and Kolss 2009 \cite{long} went further, and developed a method for long-distance word reordering, which works good on German-English translation task due to the long-distance shift of verbs. The method extracts discontinuous reordering rules in addition to the continuous ones, which contains a placeholder to match several words and enables the word to shift cross long distance.

@@

\begin{figure}
\centering
\input{extract.tikz}
\caption{A phrase with its syntax tree and alignment for rule extraction}
\label{extract}
\end{figure}

@@

Table~\ref{tenw} shows the BLUE scores, absolute improvements of BLEU scores and TER scores for configurations with different reordering methods. The table consist of $2$ sections. the first row of the top section shows results of the baseline, which used no preordering at all. In the following rows of the top section, different types of reordering rules are combined gradually, with each type per row, and the results are showed. For example, the row with \emph{$+$MLT} Rules presents the configuration with all the rule types including MLT rules and all the other rules in the rows above. All improvements are absolute improvements of BLEU scores in comparison to the baseline. Each row with a certain reordering type presents all the different variations of this type and the best score under these configurations are shown. For example, long rules include the left rules and right rules, and the tree rules include the partial rules and recursive application. The baseline used a monotone decoder and no preordering. The row with \emph{oracle reordering} shows the results from the configuration that used the oracle reordering as input. The results of oracle reordering can be used for analyzing the potential of source sentence reordering. In the lower section of the table, different rule types are not combined and the effect of each rule type is shown.

@@

\begin{table}
\centering
\STautoround*{2}
\begin{spreadtab}{{tabular}{|l|r|r|r|}}\hline
@				& @BLEU (\%) & @Imprv. & @TER (\%) \\ \hline
@Baseline		& 21.80 & & 62.09 \\ \hline
@+Short Rules	& 22.90 & :={b3 - b2}& 61.64 \\ \hline
@+Long Rules   & 23.13 & :={b4 - b2} & 61.43\\ \hline
@+Tree Rules   & 23.84 & :={b5 - b2} & 60.95\\ \hline
\textbf{@+MLT Rules}    & \textbf{:={24.14}} & \textbf{:={b6 - b2}} & \hphantom{xxx} \textbf{:={60.79}}\\ \hline
@Oracle Reordering & :={26.80} & :={b7 -b2} & :={56.97} \\ \hline
\hline
@Long Rules   & 22.10 & :={b8 - b2} & 62.21\\ \hline
@Tree Rules   & 23.35 & :={b9 - b2} & 61.52\\ \hline
\textbf{@MLT Rules}    & \textbf{:={23.96}} & \textbf{:={b10 - b2}} & \textbf{:={60.83}}\\ \hline
\end{spreadtab}
\caption{Result overview of Chinese to English systems}
\label{tzhen2}
\end{table}

@@

The results shows increasing scores as we used reordering methods from short rules, long rules, tree rules to MLT rules. And better BLEU scores were achieved as we combined the different reordering rules. The MLT rules achieved better BLEU scores and TERs in both translation directions, not only when it was used alone, but also it was added to the other reordering rules. As the MLT reordering rules were used alone, it showed an improvement of $1.61$ in BLEU score in the English-to-Chinese translation direction and an improvement of $2.16$ in BLEU score in the Chinese-to-English translation direction, in comparison with the baseline, which used no reordering at all. As the MLT reordering rules were combined with the other existing reordering rules, a further improvement of $0.43$ in BLEU score (from $13.38$ to $13.81$) was shown in the English-to-Chinese translation direction, as well as a further improvement of $0.3$ in BLEU score (from $23.84$ to $24.14$) in the Chinese-to-English translation direction.

>>>>

\address{Institute for Anthropomatics \\
Karlsruhe Institute of Technoloy, Germany \\
{\small \tt utcur@student.kit.edu, yuqi.zhang@kit.edu}
}
%
\begin{document}
\maketitle
%
\begin{abstract}
We propose a novel data-driven rule-based preordering approach, which uses the information of syntax tree and word alignment to reorder the words in source sentences before decoding in a phrase-based SMT system between English and Chinese. The preordering algorithm extracts reordering patterns from multiple levels of the syntax trees of the training data and applies the rules on the source sentences in a similar manner. We've conducted experiments in English-to-Chinese and Chinese-to-English translation directions. Our results show that the approach has led to improved translation quality both when it was applied separately on the baseline or when it was combined with some other reordering approaches. As our reordering approach was used alone, it showed an improvement of $1.61$ in BLEU score in the English-to-Chinese translation direction and an improvement of $2.16$ in BLEU score in the Chinese-to-English translation direction, in comparison with the baseline, which used no preordering at all. As our preordering approach were combined with the short rule \cite{short}, long rule \cite{long} and tree rule \cite{tree} based preordering approaches, it showed an improvement of $0.43$ in BLEU score in the English-to-Chinese translation direction and an improvement of $0.3$ in BLEU score in the Chinese-to-English translation direction. Through the translations that used our preordering approach, we've also found improvements in syntactic structure.
\end{abstract}
%
\section{Introduction}

@@

Rottmann and Vogel 2007 \cite{short} introduced the idea of extracting reordering rules from the POS tag sequences of training data and use them for reordering. Niehues and Kolss 2009 \cite{long} went further, and developed a method for long-distance word reordering, which works good on German-English translation task due to the long-distance shift of verbs. The method extracts discontinuous reordering rules in addition to the continuous ones, which contains a placeholder to match several words and enables the word to shift cross long distance.

@@

\begin{figure}
\centering
\input{extract.tikz}
\caption{A phrase with its syntax tree and word alignment for rule extraction}
\label{extract}
\end{figure}

@@

Table~\ref{tenw} shows the BLEU scores, absolute improvements of BLEU scores and TER scores for configurations with different reordering methods. The table consist of $2$ sections. the first row of the top section shows results of the baseline, which used no preordering at all. In the following rows of the top section, different types of reordering rules are combined gradually, with each type per row, and the results are showed. For example, the row with \emph{$+$MLT} Rules presents the configuration with all the rule types including MLT rules and all the other rules in the rows above. All improvements are absolute improvements of BLEU scores in comparison to the baseline. Each row with a certain reordering type presents all the different variations of this type and the best score under these configurations are shown. For example, long rules include the left rules and right rules, and the tree rules include the partial rules and recursive application. The baseline used a monotone decoder and no preordering. The row with \emph{oracle reordering} shows the results from the configuration that used the oracle reordering as input. The results of oracle reordering can be used for analyzing the potential of source sentence reordering. In the lower section of the table, different rule types are not combined and the effect of each rule type is shown.

@@

\begin{table}
\centering
\STautoround*{2}
\begin{spreadtab}{{tabular}{|l|r|r|r|}}\hline
@				& @BLEU (\%) & @Imprv. & @TER (\%) \\ \hline
@Baseline		& 21.80 & & 62.09 \\ \hline
@+Short Rules	& 22.90 & :={b3 - b2}& 61.64 \\ \hline
@+Long Rules   & 23.13 & :={b4 - b2} & 61.43\\ \hline
@+Tree Rules   & 23.84 & :={b5 - b2} & 60.95\\ \hline
\textbf{@+MLT Rules}    & \textbf{:={24.14}} & \textbf{:={b6 - b2}} & \hphantom{xxx} \textbf{:={60.79}}\\ \hline
@Oracle Reordering & :={26.80} & :={b7 -b2} & :={56.97} \\ \hline
\hline
@Long Rules   & 22.10 & :={b8 - b2} & 62.21\\ \hline
@Tree Rules   & 23.35 & :={b9 - b2} & 61.52\\ \hline
\textbf{@MLT Rules}    & \textbf{:={23.96}} & \textbf{:={b10 - b2}} & \textbf{:={60.83}}\\ \hline
\end{spreadtab}
\caption{Result overview of Chinese-to-English systems}
\label{tzhen2}
\end{table}

@@

The results shows increasing scores as we used reordering methods from short rules, long rules, tree rules to MLT rules. And better BLEU scores were achieved as we combined the different reordering rules. The MLT rules achieved better BLEU scores and TER scores in both translation directions, not only when it was used alone, but also it was added to the other reordering rules. As the MLT reordering rules were used alone, it showed an improvement of $1.61$ in BLEU score in the English-to-Chinese translation direction and an improvement of $2.16$ in BLEU score in the Chinese-to-English translation direction, in comparison with the baseline, which used no reordering at all. As the MLT reordering rules were combined with the other existing reordering rules, a further improvement of $0.43$ in BLEU score (from $13.38$ to $13.81$) was shown in the English-to-Chinese translation direction, as well as a further improvement of $0.3$ in BLEU score (from $23.84$ to $24.14$) in the Chinese-to-English translation direction.
